Visualizing Project Evolution Through Abstract Syntax Tree Analysis

### 2016-06-20 paper/document.tex
Go through 50% of the paper.

The plain text nature of software does not stop at source code level. The history and evolution of a software project is often tracked by version control systems, such as Git, in a programming language agnostic way. Updates to source code, documentation, and tests are recorded in the form of commits, which store the differences in lines of text in a file between each subsequent revision of the software. Version control systems become vital for large projects since individual changes would be impossible to track manually. Large projects have thousands of individual changes made by many different developers. With more changes to peruse, it becomes more difficult to see how a project is evolving over time since there is so much information to look through. Hosting services for Git repositories often provide some basic tools for analyzing projects over time~\cite{github-graphs,bitbucket-graphs}; however, as we will show, these tools often lack the kind of detail to gather insightful analysis. As with source code, commits operate at the text level. While commits may store individual changes as lines of code, they imply a change in the \emph{structure} of software.  By visualizing lines changed, or files changed, many current tools for software projects rarely capture the \emph{meaning} behind each commit; lines of code alone do not describe how a repository changed at a particular point in time. Thus, information regarding contributions from developers is often misleading since the number of commits and lines of code changed do not necessarily reflect meaningful changes to a code repository~\cite{robles2014}. In this paper, we will show that abstract syntax trees (ASTs) can be used to create \emph{semantically-aware} software evolution visualizations.

Consider GitHub graphs~\cite{github-graphs}: visualizations that are easily available to the many software developers using GitHub to host their software repositories. GitHub's graphs use line and commit information exclusively. In the case of the ANTRL4 project (Figure~\ref{fig:GH_code_freq}), this code frequency chart creates more questions than it answers. What happened at those big spikes of line changes in the graph? One of the spikes is so large, we had to crop the image in order to show the magnitude of the other changes. How did these large changes affect the code base? Who committed these large changes? Was it a team of contributors or one particularly prolific programmer? Plotting the number of lines changed may be misleading. A huge spike in additions may be caused by a number of reasons, such as the inclusion of automatically-generated documentation, the inclusion of an external library, or perhaps a massive refactoring effort~\cite{Hindle:2008:LCT:1370750.1370773}.

@@

We will examine research which analyses software repositories for information on language and features usage, tools for visualizing project evolution, collaboration in software repositories and other tools for understanding repositories.

@@

Some tools examine the evolution of projects to learn other information about them. Alcocer \textit{et al.}~\cite{6650523} create a Performance Evolution Blueprint, which displays the changes in performance over time as a project evolves. Ogawa and Ma created code\_swarm~\cite{codeswarm}, a generative art visualization which presents the progressive evolution of file changes made by authors in source control repository over time. Gource~\cite{gource} is a similar visualization, showing an animated force-directed tree of the file hierarchy as authors make changes to files in the source code repository over time. Ens \textit{et al.} created ChronoTwigger to visualize the co-evolution of source and test files~\cite{6980223}. This allowed testing practices to be more closely examined. Anslow \textit{et al.}~\cite{sourcevis} created SourceVis, a suite of software visualizations including visualizations for semantically-aware project evolution. SourceVis displays the evolution of Java source code metrics such as number of packages, classes, and methods, and fields.

@@

There is also research on how developers use software repositories to collaborate. Wagstrom \textit{et al.}~\cite{Patrick:Wagstrom:2012} explored the roles that developers take in a networked, social development environments like GitHub. These role were defined as their level of contribution as well as the types of issues they dealt with. They found that developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects. Lee \textit{et al.}~\cite{lee2013} developed a tool for visualizing the branching structure of Git repositories. This was used to analyze concurrent workflows in popular open source repositories. Elsen also created a visualization for branches in Git, with the ability to expand branches to view the directory structure of a project at that point~\cite{6650522}.   Shrestha \textit{et al.} made a visualization for the geographic location of a commit and~\cite{6650532} and Minelli \textit{et al.}~\cite{6980226} looked at the real time changes developers made to a repository.

@@

Some visualization tools are used for managing or making decisions about software rather than study the changes taking place. The service EasyBi~\cite{EasyBi} provides additional visualizations for Git repositories for the purpose of business decisions. EasyBi works by taking the git log file and generating charts and graphs from it. It tries to solve the issue of summarizing many commits and has greater controls for granularity and authors but still operates on commits and line differences. The Gitinspector project~\cite{Gitinspector} is an open source tool for the statistical analysis of Git repositories. It provides an advanced timeline view of changes per author and cumulative statistics but again relies upon the line differences to summarize the work of each author.

Developers using the massively popular GitHub website for their Git repository hosting instantly have access to basic history visualizations, known as GitHub Graphs (see Figure~\ref{fig:GH_code_freq})~\cite{github-graphs}. GitHub Graphs tracks contributions based only on the number of commits and number of lines changed. This has the advantage of not relying on the of code which was committed. It does not matter what language is used or if errors exist in the commit, GitHub will simply output the number of line additions and deletions. The issue with this type of analysis is that it does not give a clear idea of what is being changed, only that changes have occurred. This data may be lacking, or even misleading which contributes to the need for better tools.

@@

Given a Git repository, we want to extract per-author type changes over time. We accomplish this by computing the differences between abstract syntax trees of every consecutive pair of commits (Figure~\ref{fig:context}). This extracted data can then be used in a variety of visualizations as discussed in Section~\ref{sec:viz}. The reason for using ASTs is that they capture the semantics of the code far more than simple line counts.

An \emph{abstract syntax tree} (AST) is a data structure derived from source code that breaks its syntactic constructs into a tree. Each node in the tree represents a syntactically valid chunk of code. This captures the structure of the code in an abstract and tractable manner. When a programmer makes a non-trivial change in code in a file, that difference will be reflected in the AST\@. Non-trivial changes are those that affect how a compiler will interpret the source code, and may have effects on the executable code. Trivial changes are those that have no affect on how a compiler will interpret the source text, such as changing insignificant whitespace or comments in source code. Thus, ASTs of consecutive revisions in a code repository can be compared to see what \emph{structures} a programmer is modifying.

@@

Each change of declaration or invocation is detected and tallied per file per commit. These changes are emitted in line-based file format. For example, the following is the output of a commit that changed two declaration of \texttt{String} variables to \texttt{URI} variables:

@@

In order to gather all the changes for a given repository over time, we generated ASTs for each commit and compared the difference. We did this by running \texttt{git log} over the Git repository and listing all the commit IDs in chronological order (Figure~\ref{fig:commits}). For each commit, we collected the following information, as can be seen in Figure~\ref{fig:data-extracted}: the author's name, date of commit, commit ID (SHA), commit message, files the commit edited, and files in the repository at the time of the commit. For all the edited Java files we also compared the file at the time of the commit with its parent commit using AST differencing as described above. We achieved this by invoking \texttt{git difftool} and specifying the tool to be used as our AST differencing tool. One case where \texttt{git difftool} fails is when the commit has no parent, such as the initial commit. We were able to handle this by calling \texttt{git show} and saving the output to a temporary file to be used with the AST differencing tool. When calling \texttt{git show} instead of \texttt{git difftool} we treat everything as an addition, as in commit A in Figure~\ref{fig:commits}. Since we compare each commit to its parent as reported by Git, we are able to account for any branching that happened outside the primary branch of the repository (usually named \texttt{master}). If a commit has more than one parent, like commit D in Figure~\ref{fig:commits}, it indicates a merge commit in Git.

We deliberately chose to ignore merge commits. A merge commit has two cases to consider: The first case is when the parents of the commit can be automatically merged. In this case all the relevant changes are already recorded in the parent of the commits so we can safely ignore the merge commit~\cite{git:merge}. The second case occurs when the parents cannot be merged automatically: There is a merge conflict which requires manual resolution on the part of a developer. In this case the Git user will try to resolve the issue by modifying code. If we decide to include the merge commit we run into the issue of to how to compare the merge commit with its multiple parent commits and the possible modifications made to resolve the merge conflict. We could consider comparing the merge commit with only one parent commit at a time, however; we will end up counting all the changes already recorded in the other parent branch. Additionally, all changes made during the resolution of a conflict will be attributed to the author making the merge; this may exclude the author that actually wrote the code in the first place. The other option is to simply ignore the merge commit such that we do not count the changes done during a merge commit. We made the assumptions that merge conflicts occur infrequently and that changes done to resolve a merge conflict are usually minor. Therefore, we chose to ignore merge commits. Note that fast-forwards are not \emph{true merges}, hence they do not create merge commits~\cite{git:merge}; none of the previous cases involving merges occur.

@@

Although rich in detail, the AST diff data is difficult to analyze if displayed in its raw form; hence, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}). The source code for TypeV---both the AST differencing component and the web application--- is available online on GitHub.\footnote{\url{https://github.com/mdfeist/TypeV}} A screencast demonstrating its use is also available online.\footnote{\url{https://archive.org/details/typev-demo}}

@@

To allow for a more detailed analysis of a specific component of the repository we added filtering tools. Users can filter by type name, view changes that fall within a specific start and end date, and select whether to see only declarations of types, types used in declarations and invocations, or only invocations.

@@

The purpose of displaying stacked bar charts is that they are a familiar visualization. They can be viewed as an enhanced version of GitHub's code frequency graph~\cite{github-graphs}, with semantic data. We chose to create a similar visualization to GitHub to work with a visualization people may already be familiar with and evaluate our visualization tools' density of data.

The stacked bar graphs succinctly display the relative amounts of additions and deletions per each time period. This creates a dense view of the project's evolution. We chose to make the height of the bars relative to only the type itself to make it easy for a user to scan a single type and see when it was used modified. This makes tasks such finding a refactoring (see \ref{sec:antlr4}) very clear to see simply by browsing the line. We stacked the additions on top of the deletions such that users can immediately compare the relative frequency of changes of a single type throughout the chosen time span.

Our choice of colours was motivated to be high contrast and culturally indicative of their meaning. The particular colours chosen are high contrast, and are distinct to those experiencing protanomaly or deuteranomaly---red-green colour blindness. The red colour suggests a negative valence~\cite{moller2009}, hence suggesting deletions. The blue serves as a contrast to the red, indicating the opposite action (in this case, additions). Since

The rows are sorted so that the graph for the most used types or methods appear the top and are in descending order of use such that the most frequently used type is the first visible. The height allows users to see when certain types are being edited and ensures that highly used types do not overpower less used type making their changes less significant.

@@

To determine if type coverage yields different information than file coverage, we analyzed four open source Java projects: ANTLR4, Apache BookKeeper, Apache Curator and Apache Tika. Three of the four projects were collected from the Apache foundation, as they host a large number of well-curated open source Java projects. A summary of the projects studied is given in Table~\ref{tab:project-summary}. We counted the file coverage and type coverage on a commit-by-commit basis for each author. That is, each datum represents a commit with three dimensions: file coverage, type coverage, and its commit date. Data is sorted in ascending order of commit date. Table~\ref{tab:summary} lists a summary of the statistics measured. In general, type coverage and file coverage are strongly correlated, both using Pearson (linear) correlation, and Spearman (rank) correlation. As type coverage and file coverage are both cumulative measures, positive correlation is expected. We displayed the density distributions for two different projects: ANTLR4 (Figure~\ref{fig:antlr4}) and Apache Tika (Figure~\ref{fig:tika-density}). However, even when the distributions are significantly different (as with ANTLR and Curator), the correlation is still strongly positive, but not 100\%. This means that when the coverage for types and files of a commit increase, the relationship is not necessarily one-to-one. This is remarkable in a language like Java, as classes and files are generally in a one-to-one relationship.

@@

Using TypeV we can can see that at certain times there were major changes across all types. This shows us when major changes to the code were being done versus code maintenance.

@@

The purpose of \emph{deprecating} software features is to give developers time to change projects to follow a new or better standard without breaking the code. When a piece of code becomes deprecated, continuing its use may lead to design flaws or even security vulnerabilities. For example, in Apache Nifi, an earlier version of the software used a compromised key-derivation function; hence, this class was deprecated, but left in for compatibility with older versions of the software~\cite{nifi}. Viewing changes in terms of types can show when usage of deprecated classes are added or removed from a project as well as how many remain over time. Given the time at which a feature was deprecated, seeing the progress of replacement has some important implications. In the case of a security flaw, that time could indicated times at which systems or users were vulnerable. If a feature is marked for removal, a project using a deprecated type has a limited amount of time to make the changes. Removing all instances of deprecated features could also mean a project can be upgraded to use new tools or features that were not possible while the deprecated features were present.

>>>>

The plain text nature of software does not stop at the source code level. The history and evolution of a software project is often tracked by version control systems, such as Git, in a programming language agnostic way. Updates to source code, documentation, and tests are recorded in the form of commits, which store the differences in lines of text in a file between each subsequent revision of the software. Version control systems become vital for large projects since individual changes would be impossible to track manually. Large projects have thousands of individual changes made by many different developers. With more changes to peruse, it becomes more difficult to see how a project is evolving over time since there is so much information to look through. Hosting services for Git repositories often provide some basic tools for analyzing projects over time~\cite{github-graphs,bitbucket-graphs}; however, as we will show, these tools often lack the kind of detail to gather insightful analysis. As with source code, commits operate at the text level. While commits may store individual changes as lines of code, they imply a change in the \emph{structure} of software.  By visualizing lines changed, or files changed, many current tools for software projects rarely capture the \emph{meaning} behind each commit; lines of code alone do not describe how a repository changed at a particular point in time. Thus, information regarding contributions from developers is often misleading since the number of commits and lines of code changed do not necessarily reflect meaningful changes to a code repository~\cite{robles2014}. In this paper, we will show that abstract syntax trees (ASTs) can be used to create \emph{semantically-aware} software evolution visualizations.

Consider GitHub graphs~\cite{github-graphs}: visualizations that are easily available to software developers using GitHub to host their Git software repositories. GitHub's graphs use line and commit information exclusively. In the case of the ANTLR4 project (Figure~\ref{fig:GH_code_freq}), this code frequency chart creates more questions than it answers. What happened at those big spikes of line changes in the graph? One of the spikes is so large, we had to crop the image in order to show the magnitude of the other changes. How did these large changes affect the code base? Who committed these large changes? Was it a team of contributors or one particularly prolific programmer? Plotting the number of lines changed may be misleading. A huge spike in additions may be caused by a number of reasons, such as the inclusion of automatically-generated documentation, the inclusion of an external library, or perhaps a massive refactoring effort~\cite{Hindle:2008:LCT:1370750.1370773}.

@@

We examined research which analyses software repositories for information on language and feature usage, tools for visualizing project evolution, collaboration in software repositories and other tools for understanding repositories.

@@

Some tools examine the evolution of projects to learn other information about them. Alcocer \textit{et al.}~\cite{6650523} create a Performance Evolution Blueprint, which displays the changes in performance over time as a project evolves. Ogawa and Ma created code\_swarm~\cite{codeswarm}, a generative art visualization which presents the progressive evolution of file changes made by authors in a source control repository over time. Gource~\cite{gource} is a similar visualization, showing an animated force-directed tree of the file hierarchy as authors make changes to files in the source code repository over time. Ens \textit{et al.} created ChronoTwigger to visualize the co-evolution of source and test files~\cite{6980223}. This allowed testing practices to be more closely examined. Anslow \textit{et al.}~\cite{sourcevis} created SourceVis, a suite of software visualizations including visualizations for semantically-aware project evolution. SourceVis displays the evolution of Java source code metrics such as number of packages, classes, methods, and fields.

@@

There is also research on how developers use software repositories to collaborate. Wagstrom \textit{et al.}~\cite{Patrick:Wagstrom:2012} explored the roles that developers take in networked, social development environments like GitHub. These roles were defined by their level of contribution as well as the types of issues they dealt with. They found that developers will take on multiple roles in a project and will sometimes fulfill the same role across different projects. Lee \textit{et al.}~\cite{lee2013} developed a tool for visualizing the branching structure of Git repositories. This was used to analyze concurrent workflows in popular open source repositories. Elsen also created a visualization for branches in Git, with the ability to expand branches to view the directory structure of a project at that point~\cite{6650522}.   Shrestha \textit{et al.} made a visualization for the geographic location of a commit and~\cite{6650532} and Minelli \textit{et al.}~\cite{6980226} looked at the real time changes developers made to a repository.

@@

Some visualization tools are made for managing or making decisions about software rather than studying the changes taking place in an academic context. EasyBi~\cite{EasyBi} is a service that provides additional visualizations for Git repositories for the purpose of business decisions. EasyBi works by taking the \texttt{git log} file and generating charts and graphs from it. It tries to solve the issue of summarizing many commits and has greater controls for granularity and authors but still operates on commits and line differences. The Gitinspector project~\cite{Gitinspector} is an open source tool for the statistical analysis of Git repositories. It provides an advanced timeline view of changes per author and cumulative statistics but again relies upon the line differences to summarize the work of each author.

Developers using the massively popular GitHub website for hosting their Git repositories instantly have access to basic visualizations, known as GitHub Graphs (see Figure~\ref{fig:GH_code_freq})~\cite{github-graphs}. GitHub Graphs tracks contributions based only on the number of commits and number of lines changed. This has the advantage of not relying on the code which was committed. It does not matter what language is used or if errors exist in the commit---GitHub will simply output the number of line additions and deletions. The issue with this type of analysis is that it does not give a clear idea of what is being changed, only that changes have occurred. This data may be lacking, or even misleading which contributes to the need for better tools.

@@

Given a Git repository, we want to extract per-author type changes over time. We accomplish this by computing the differences between abstract syntax trees of every consecutive pair of commits (Figure~\ref{fig:context}). This extracted data can then be used in a variety of visualizations as discussed in Section~\ref{sec:viz}. The reason for using ASTs is that they capture the semantics of the code far better than simple line counts.

An \emph{abstract syntax tree} (AST) is a data structure derived from source code that breaks its syntactic constructs into a tree. Each node in the tree represents a syntactically valid chunk of code. This captures the structure of the code in an abstract and tractable manner. When a programmer makes a non-trivial change in code in a file, that difference will be reflected in the AST\@. Non-trivial changes are those that affect how a compiler will interpret the source code, and may have effects on the executable code. Trivial changes are those that have no effect on how a compiler will interpret the source text, such as changing insignificant whitespace or comments in source code. Thus, ASTs of consecutive revisions in a code repository can be compared to see what \emph{structures} a programmer is modifying.

@@

Each change of declaration or invocation is detected and tallied per file per commit. These changes are emitted in an ad hoc file format. For example, the following is the output of a commit that changed two declaration of \texttt{String} variables to \texttt{URI} variables:

@@

In order to gather all the changes for a given repository over time, we generated ASTs for each commit and compared the difference. We did this by running \texttt{git log} over the Git repository and listed all commit IDs in chronological order (Figure~\ref{fig:commits}). For each commit, we collected the following information, as can be seen in Figure~\ref{fig:data-extracted}: the author's name, date of commit, commit ID (SHA), commit message, files the commit edited, and files in the repository at the time of the commit. For all the edited Java files we also compared the file at the time of the commit with its parent commit using AST differencing as described above. We achieved this by invoking \texttt{git difftool} and specifying our AST differencing tool. One case where \texttt{git difftool} fails is when the commit has no parent, such as the initial commit. We were able to handle this by calling \texttt{git show} and saving the output to a temporary file to be used with the AST differencing tool. When calling \texttt{git show} instead of \texttt{git difftool} we treat everything as an addition, as in commit A in Figure~\ref{fig:commits}. Since we compare each commit to its parent as reported by Git, we are able to account for any branching that happened outside the primary branch of the repository (usually named \texttt{master}). If a commit has more than one parent, like commit D in Figure~\ref{fig:commits}, it indicates a merge commit in Git.

We deliberately chose to ignore merge commits. A merge commit has two cases to consider: The first case is when the parents of the commit can be automatically merged. In this case all the relevant changes are already recorded in the parent of the commit so we can safely ignore the merge commit~\cite{git:merge}. The second case occurs when the parents cannot be merged automatically: this is a \emph{merge conflict} which requires manual resolution on the part of a developer. In this case, the developer will try to resolve the issue by modifying code. If we decide to include the merge commit we run into several edge case issues such as how to compare the merge commit with its multiple parent commits and how to account for the possible modifications made to resolve the merge conflict. We could consider comparing the merge commit with only one parent commit at a time; however, we will end up counting all the changes already recorded in the other parent branch. Additionally, all changes made during the resolution of a conflict will be attributed to the author making the merge. This may exclude the author that actually wrote the code in the first place. The other option is to simply ignore the merge commit such that we do not count the changes done during a merge commit. We made the assumptions that merge conflicts occur infrequently and when they do occur, the changes made to resolve a merge conflict are usually minor. Therefore, we chose to ignore merge commits. Note that fast-forwards are not \emph{true merges}, hence they do not create merge commits~\cite{git:merge} so none of the previous cases involving merges apply.

@@

Although rich in detail, the AST diff data is difficult to analyze if displayed in its raw form; hence, we created an HTML5 web visualization application which is intended for developers, project managers, and researchers to use (Figure~\ref{fig:main-view}). The source code for TypeV---both the AST differencing component and the web application---is available online on GitHub.\footnote{\url{https://github.com/mdfeist/TypeV}} A screencast demonstrating its use is also available online.\footnote{\url{https://archive.org/details/typev-demo}}

@@

To allow for a more detailed analysis of a specific component of the repository we added filtering tools. Users can filter by type name; view changes that fall within a specific start and end date; and select whether to see only declarations of types, types used in declarations and invocations, or only invocations.

@@

The purpose of displaying stacked bar charts is that they are a familiar visualization. They can be viewed as an enhanced version of GitHub's code frequency graph~\cite{github-graphs}, with semantic data. We chose to create a similar visualization to GitHub to work with a layout people may already be familiar with and to evaluate our visualization tools' density of data.

The stacked bar graphs succinctly displays the relative amounts of additions and deletions per each time period. This creates a dense view of the project's evolution. We chose to make the height of the bars relative to only the type itself to make it easy for a user to scan a single type and see when it was modified. This makes tasks such as finding a refactoring (see Section~\ref{sec:antlr4}) very clear to see simply by browsing the line. We stacked the additions on top of the deletions such that users can immediately compare the relative frequency of changes of a single type throughout the chosen time span.

Our choice of colours was motivated to be high contrast and culturally indicative of their meaning. The particular colours chosen are high contrast, and are distinct to those experiencing protanomaly or deuteranomaly---red-green colour blindness. The red colour suggests a negative valence~\cite{moller2009}, hence suggesting deletions. The blue serves as a contrast to the red, indicating the opposite action (in this case, additions).

The rows are sorted so that the graph for the most used types or methods appear the top and are in descending order of use such that the most frequently used type is the first visible. The height allows users to see when certain types are being edited and ensures that highly used types do not overpower less used types making their changes less significant.

@@

To determine if type coverage yields different information than file coverage, we analyzed four open source Java projects: ANTLR4, Apache BookKeeper, Apache Curator and Apache Tika. Three of the four projects were collected from the Apache foundation, as they host a large number of well-curated open source Java projects. A summary of the projects studied is given in Table~\ref{tab:project-summary}. We counted the file coverage and type coverage on a commit-by-commit basis for each author. That is, each datum represents a commit with three dimensions: file coverage, type coverage, and its commit date. Data is sorted in ascending order of commit date. Table~\ref{tab:summary} lists a summary of the statistics measured.

We displayed the density distributions for two different projects: ANTLR4 (Figure~\ref{fig:antlr4-density}) and Apache Tika (Figure~\ref{fig:tika-density}). These plots show the distribution of the amount of commits that have X\% of type and file coverage. For example, Figure~\ref{fig:tika-density} shows that there are many commits with 50\% type coverage and very few commits with 50\% file coverage. This shows that even though the type and file coverage plots may have a similar shape the commits are showing different levels of information.

In general, type coverage and file coverage are strongly correlated, both using Pearson (linear) correlation, and Spearman (rank) correlation. As type coverage and file coverage are both cumulative measures, positive correlation is expected. However, even when the distributions are significantly different (as with ANTLR and Curator), the correlation is still strongly positive, but not 100\%. This means that when the coverage for types and files of a commit increase, the relationship is not necessarily one-to-one. This is remarkable in a language like Java, as classes and files are generally in a one-to-one relationship.

@@

Using TypeV we can see that at certain times there were major changes across all types. This shows us when major changes to the code were being done versus code maintenance.

@@

The purpose of \emph{deprecating} software features is to give developers time to change projects to follow a new or better standard without breaking existing code. When a piece of code becomes deprecated, continuing its use may lead to design flaws or even security vulnerabilities. For example, in Apache Nifi, an earlier version of the software used a compromised key-derivation function; hence, this class was deprecated, but left in for compatibility with older versions of the software~\cite{nifi}. Viewing changes in terms of types can show when usage of deprecated classes are added or removed from a project as well as how many remain over time. Given the time at which a feature was deprecated, seeing the progress of replacement has some important implications. In the case of a security flaw, that time could indicate times at which systems or its users were vulnerable. If a feature is marked for removal, a third party using the deprecated feature has a limited time to make the appropriate changes. Removing all instances of deprecated features could also mean a project can be upgraded to use new tools or features that were not possible when the deprecated features were present.
